Effects of Reynolds Number on the Performance and Near-Wake of a Cross-Flow Turbine

### 2014-06-22 paper.tex
Change wake measurement grid figure to newer drawing
Add intro comments about Reynolds number dependence and CAD model reference

@@

Experiments were performed in the University of New Hampshire's tow/wave tank
turbine test bed. The turbine model is dubbed the UNH-RVAT, for Reference
Vertical Axis Turbine, which is designed to be a generic case for numerical
model testing, similar to the Sandia National Labs/US Department of Energy RM2
River Turbine. Details of the turbine and experimental setup are described in
\cite{Bachant2014_JOT}.

\todo[inline]{Put up CAD model for citation/download on figshare.}

@@

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{figures/wake_meas_grid}
\caption{Wake measurement coordinate system and locations.}
\label{fig:wake_locations}
\end{figure}

>>>>

\subsection{Previous Work}
\todo[inline]{Add lit review here}

\subsection{Causes for Reynolds Number Dependence}

Turbine performance is strongly dependent on the lift-to-drag ratio $l/d$ acting
on the blades, which is a function of the blades' angles of attack $\alpha$.
These angles of attack are a function of turbine tip speed ratio, induction
(slowing/turning of the free stream flow prior to reaching the turbine), and
turbine rotation or azimuthal angle $\theta$. Looking at static airfoil data, we
see that the static stall angle---the angle just beyond that at which $l/d$ is
maximum, characterized by flow separation near the leading edge---increases with
blade chord Reynolds number $Re_c$. This can be explained as the transition to a
turbulent boundary layer adjacent to the blade delaying separation. However,
static foil performance does not tell the whole story for a cross-flow turbine.
The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
turbine implies the occurrence of dynamic stall at tip speed ratios at and below
those of maximum rotor torque\cite{Para2002}. We then must understand the
effects of Reynolds number on the dynamic stall process.

As previously mentioned, the near-wake of a CFT is not very similar to a simple
drag source, or plane shear flow. One of the mean wake's streamwise vortex pairs
can be predicted by potential flow theory as a result of finite-span blades
producing lift (blade end effects), but the other---caused by viscous effects,
i.e., dynamic stall---is not. We can attempt to predict changes
\todo[inline]{Finish this thought with Reynolds number dependence of these
simpler flows, and set up the question whether or not we can assume these will
combine linearly to produce the wake $Re$-dependence we see.}

@@

Experiments were performed in the University of New Hampshire's tow/wave tank
turbine test bed. The turbine model is dubbed the UNH-RVAT, for Reference
Vertical Axis Turbine, which is designed to be a generic case for numerical
model testing, similar to the Sandia National Labs/US Department of Energy RM2
River Turbine. Details of the turbine and experimental setup are described in
\cite{Bachant2014_JOT}, and a CAD model of the turbine is available from
\cite{Bachant2014_CAD}.

@@

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{figures/turbine_coordinate_system}
\caption{Wake measurement coordinate system and locations.}
\label{fig:wake_locations}
\end{figure}
